### 2018-05-02 Maths Notes/Mechanics/M2/Statics of rigid bodies/Statics of rigid bodies.tex
Moment clarification

$$\textrm{Moment of a force about a point P}=|\textrm{Force}|\times\textrm{Perpendicular distance of the line of action of the force from the point P}$$
Essentially meaning, when taking non linear moments, the distances used should be the shortest distance from the point to where the force is acting.
\newpage
\begin{center}
\underline{\Huge Equilibrium of rigid bodies example}
\end{center}
\includegraphics[width=8cm]{eq_rigid.png}\\
\textit{The diagram above shows a uniform rod AB of mass m and length 4a. The end A of the rod is
freely hinged to a point on a vertical wall. A particle of mass m is attached to the rod at B. One
end of a light inextensible string is attached to the rod at C, where AC = 3a. The other end of the
string is attached to the wall at D, where AD = 2a and D is vertically above A. The rod rests
horizontally in equilibrium in a vertical plane perpendicular to the wall and the tension in the
string is T. }\\
\\
\textit{Show that T = $mg\sqrt{13}$}\\
\\
Take moments about A 
$$3a\times T\cos\theta=2amg+4amg$$
Calculate $\cos\theta$ from the lengths on the diagram
$$\cos\theta=\frac{2}{\sqrt{3^2+2^2}}=\frac{2}{\sqrt{13}}$$
Substitute in the value of $\cos\theta$ and divide through by $a$
$$\frac{6}{\sqrt{13}}T=6mg$$
Multiply both sides by $\frac{\sqrt{13}}{6}$
$$T=mg\sqrt{13}$$
\\
\textit{The particle of mass m at B is removed from the rod and replaced by a particle of mass M which
is attached to the rod at B. The string breaks if the tension exceeds $2mg\sqrt{13}$. Given that the
string does not break,\\
show that $M\leqslant\frac{5}{2}m$}\\
\\
Rewrite the moments equation with the new information from the question
$$3a\times T\cos\theta=2amg+4aMg$$
Write the inequality given
$$T\leqslant 2mg\sqrt{13}$$
Substitute in the value for T
$$\frac{2mg+4Mg}{6}\sqrt{13}\leqslant2mg\sqrt{13}$$
Simplify
$$mg+2Mg\leqslant 6mg$$
$$M\leqslant\frac{5}{2}m$$
\end{document}

>>>>

$$\textrm{Moment of a force about a point P}=|\textrm{Force}|\times\textrm{Perpendicular distance of the line of action of the force from the point P}$$
Essentially meaning, when taking non linear moments, the distances used should be the shortest distance from the point to where the force is acting.\\
The line of action is the line, both backwards and forwards from where the force is acting.
\newpage
\begin{center}
\underline{\Huge Equilibrium of rigid bodies example}
\end{center}
\includegraphics[width=8cm]{eq_rigid.png}\\
\textit{The diagram above shows a uniform rod AB of mass m and length 4a. The end A of the rod is
freely hinged to a point on a vertical wall. A particle of mass m is attached to the rod at B. One
end of a light inextensible string is attached to the rod at C, where AC = 3a. The other end of the
string is attached to the wall at D, where AD = 2a and D is vertically above A. The rod rests
horizontally in equilibrium in a vertical plane perpendicular to the wall and the tension in the
string is T. }\\
\\
\textit{Show that T = $mg\sqrt{13}$}\\
\\
Take moments about A 
$$3a\times T\cos\theta=2amg+4amg$$
Calculate $\cos\theta$ from the lengths on the diagram
$$\cos\theta=\frac{2}{\sqrt{3^2+2^2}}=\frac{2}{\sqrt{13}}$$
Substitute in the value of $\cos\theta$ and divide through by $a$
$$\frac{6}{\sqrt{13}}T=6mg$$
Multiply both sides by $\frac{\sqrt{13}}{6}$
$$T=mg\sqrt{13}$$
\\
\textit{The particle of mass m at B is removed from the rod and replaced by a particle of mass M which
is attached to the rod at B. The string breaks if the tension exceeds $2mg\sqrt{13}$. Given that the
string does not break,\\
show that $M\leqslant\frac{5}{2}m$}\\
\\
Rewrite the moments equation with the new information from the question
$$3a\times T\cos\theta=2amg+4aMg$$
Write the inequality given
$$T\leqslant 2mg\sqrt{13}$$
Substitute in the value for T
$$\frac{2mg+4Mg}{6}\sqrt{13}\leqslant2mg\sqrt{13}$$
Simplify
$$mg+2Mg\leqslant 6mg$$
$$M\leqslant\frac{5}{2}m$$
\end{document}
